\section*{Exercise 3}

True or False? For every $\varepsilon > 0$, there exists an NP - complete problem that can be solved deterministicly in
$O(2^{n^\varepsilon})$ steps. Explain your answer.

\textbf{Solution:}\\

We proof this part by construction a language for each $\varepsilon > 0$, that is $NP - complete$ and can be solved in
the given time.\\

Let $PAD-SAT_\varepsilon = \{ (\Psi , 1^{|\Psi|^{\left\lceil \frac{k}{\varepsilon} \right\rceil}}) \; | \; \Psi \text{satisfyable in 3 cnf} \}$
where $k > 1$ is a arbitrary but fixed konstant.\\

\pagebreak

\begin{lemma}\label{ex1:t3:npc}
    $PAD-SAT_\varepsilon$ is NP-complete.
\end{lemma}

\textbf{Proof \ref{ex1:t3:npc}.}\\
This part is straight forward.\\
Given an variable assignment we can check, whether the formula $\Psi$ is satisfied in linear time, by simply iteration over
the clauses. Because the number of variables has to be strictly smaller than the formulas the assignment is polynomial bounded
in the inputsize.\\

We can next reduce $PAD-SAT_\varepsilon$ to $3-SAT$. The reduction copys $\Psi$ and throws away the rest. If 
the TM for $3 - SAT$ accepts, the word is in the language, because $\Psi$ is satisfiable. If it rejects,
$\Psi$ is not satisfiable, thus the word can't be in $PAD-SAT_\varepsilon$.

\begin{lemma}\label{ex1:t3:time}
    There exists a DTM M such that $T_M(n) \in DTIME (2^{n^\varepsilon})$.
\end{lemma}

\textbf{Proof \ref{ex1:t3:time}.}\\

From the definition we can conclude, that $|\Psi| \leq n^{\frac{\varepsilon}{k}}$ holds,
because the rest of the length of the word is filled with ones.\\

$\Psi$ is a boolean formula and this formula contains each occurence of its variables. Therefore there can't
be no more variables, than the length of $\Psi$.\\
So we can compute all Assignments, which are less than $2^{n^{\frac{\varepsilon}{k}}}$.
As said before we can iterate over all clauses and literals and check whether there occures at least one 
true in each clause. Therefore we have linear costs in the length of $\Psi$.\\

We now construct a DTM $M$.\\
Given a word $(\Psi, 1^{|\Psi|^{\left\lceil \frac{k}{\varepsilon} \right\rceil }})$, we check first
whether the number of ones is right in the end. This takes us at most $n$ time.
Next we test all assignments, if they satisfy the formula.\\

Testing takes $n^\frac{\varepsilon}{k}$ time and we have $2^{n^\frac{\varepsilon}{k}}$ possibility, so
the runtime of $M$ is $T_M (n) = n^\frac{\varepsilon}{k} 2^{n^\frac{\varepsilon}{k}}$. We ignore the
linear Faktor, because we know that we can choose a faktor so that this function is strictly greate after a given $n_0$.\\

Now we have to show, that $T_M(n) \in O(2^{n^\varepsilon})$. We do this by checking the limiting value
$$
    \underset{n \rightarrow \infty}{\lim} \frac{T_M (n)}{2^{n^\varepsilon}}
        = \underset{n \rightarrow \infty}{\lim} n^\frac{\varepsilon}{k} 2^{n^\frac{\varepsilon}{k} - n^\varepsilon}\\
        = 0
$$

Because $n^\frac{\varepsilon}{k} < n^\varepsilon$ holds for $k > 1$ and than we have a exponential function with a negative exponent
and a polynomial faktor. By knowledge of Analysis I this sequence converges to 0.\\
\mbox{} \hfill $\square$
