\subsection*{The Proof}

In this part we will show, that the follwoing holds\\

\begin{lemma}\label{multcut:runtime}\mdseries\itshape
The algorithm $\mathcal{A}$ has a runtime in $P$.
\end{lemma}
\begin{lemma}\label{multcut:apx}\mdseries\itshape
The algorithm $\mathcal{A}$ is a 2-approximation-algorithm
for the multicut problem in trees.
\end{lemma}

\begin{description}
   \item{\bfseries\itshape~Proof ~\ref{multcut:runtime}:}\\
      We see, that we have at most $k$ iterations of the while loop,
      because we cut every iteration at least one pair.\\
      In each iteration we first search the maximal depth, which we can
      obtain in $k \log \, k$ overall iterations (MaxHeap).\\
      We can optain the $e_t$ in P time. The path from $s_i$ to $t_i$
      is for every pair at most $n = |V|$ (-1). This is the amount
      of elements we have to search for the minimum next.\\
      For each of these $n$ elements we have to find all Path $P_j$
      that contains the observed edge. This can be done in ineffciently in
      $k \cdot n$ for every edge (k Paths maximal and n elements in each).
      So we find the $e_t$ in $O(n^2 k)$. $\Delta$ can be found in less
      time as easily seen. The rest of the actions needs constant time,
      or at least less than the given.\\
      So the first loop has runtime $O(n^2k^2)$. The next loop checks, for
      pair, whether we can delete one of the at most $k$ taken edges.\\
      We can loop at each path in $O(n)$ of $k$ paths and this at most $k$ times
      which leads us to $O(nk^2)$.\\
      So this algorithm runs in $O(n^2k^2)$ which obviosly lays in P.

   \item{\bfseries\itshape~Proof ~\ref{multcut:apx}:}\\
      At first we observe, that $F$ is feasable.\\
      In line 5 of the algorithm, the loop terminates, only
      if there exists no pair of vertices in the set $I$.\\
      In the loop body we remove an index only if we met one
      edge with equality in the dual. Due to complemantry slagness rule
      we know, that we have the edge in the Primal and so we removed an
      edge from the unique way from a the pair $i$.\\
      We conclude, that the set $F$ was feasable after the first loop.
      The seconde loop, will only remove an edge, if the set still remains feasable.\\
      So the resulting set $R$ must be feasable.\\

      To proof the approximation we will show the following
      \begin{lemma}\label{multcut:lcaedge}\mdseries\itshape
         For any pair $y_i > 0$ $1\leq i \leq k$ $|F \cap P_i| \leq 2$
         that is for any non-zero flow there are at most 2 edges in $F$ on the flow-path.
      \end{lemma}

      \begin{lemma}\label{multcut:ipres}\mdseries\itshape
         For $F$  $\sum_{e \in F} c_e \leq 2OPT$ holds. 
      \end{lemma}

      \begin{description}
         \item{\bfseries Proof ~\ref{multcut:lcaedge}:}
            Suppose $y_i$ set to nonzero while we consider 
            $v := lca(s_i,t_i)$. Therefore no edge of $P_i$ has been cut before.
	         Any $P_j$ $j \neq i$ considered afterwards has its $lca(s_j,t_j)$  
            in the same depth or lower than $v$. 
            Therefore if $P_j$ shares an edge with $P_i$ $lca(s_j,t_j)$ is an ancestor of $v$. 
            Thus if we cut a new edge for $P_j$ which is on	$P_i$ the Algorithm removes the lower edges, 
            since they are not neccessary for any previous $(s_k,t_k)$ the upper one
            could not be removed, because it will leave the $P_j$ without a cut and 			
            $(s_i,t_i)$ will be cut by the new edge too. Hence follows the claim.

         \item{\bfseries Proof ~\ref{multcut:ipres}:}
	         $$ \sum_{e \in F} =\sum_{e \in F} \sum_{i : e \in P_i} 
            y_i = \sum_{1 \leq i \leq k} |P_i \cap F| y_i \leq 2 \sum_{1 \leq i \leq k} y_i \leq 2OPT$$
            This follows by construction, Claim ~\ref{multcut:lcaedge} and duality.
      \end{description}
\end{description}
